ComparisonOfSeveralSequentialAndParallelDerivative-freeGlobalOptimizationAlgorithms

### 2019-01-23 2019/PAVT/pavt_paper.tex
PAVT: fix style issues

\documentclass{svproc}
%
% RECOMMENDED
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%
\usepackage{graphicx}
\usepackage{marvosym}
\usepackage{amssymb}
\usepackage{cite}
\usepackage{subfig}
% to typeset URLs, URIs, and DOIs
\usepackage{url}
\usepackage{hyperref}
\def\UrlFont{\rmfamily}

@@

\begin{document}
\mainmatter              % start of a contribution
%
\title{Comparison of several sequential and parallel derivative-free global optimization algorithms}
%
\titlerunning{Comparison of several optimization methods}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Vladislav Sovrasov\letter \and Semen Bevzuk}
%
\authorrunning{Vladislav Sovrasov et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Vladislav Sovrasov and Semen Bevzuk}
%
\institute{Lobachevsky State University of Nizhni Novgorod, Russia \\
  \email{sovrasov.vlad@gmail.com, semen.bevzuk@gmail.com}
}

@@

Nonlinear global optimization of non-convex functions is considered to be one of the most
difficult problems of mathematical programming. The finding of the global minimum of
a function of several variables often appears to be more difficult than local optimization in a
thousand-dimensional space. In local optimization applying of the simplest gradient descent
method may appear to be enough, whereas in order to \textit{guarantee} finding the global
optimum, an optimization method has to accumulate information about behavior of the
objective function within the whole search domain
\cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}. Nowadays, various stochastic global
optimization algorithms have attracted much attention, first of all, the evolution ones
\cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}. They have rather simple structure
and allow solving the problems of large dimensionality. However, the evolution algorithms ensure
the global convergence in the probabilistic sense only.

@@

Earlier, the comparison of the stochastic global optimization algorithms \cite{Ali2005, JSSv060i06}
as well as of the deterministic ones \cite{posik2012, KVASOV2018245, Liberti2005} between each
other has been considered in the literature. In these works, most of modern methods have
been studied in details, but the emphasis was made mainly on the sequential algorithms while the
reliability and the convergence speed were the main criteria of efficiency of the optimization
methods. In the majority of works, the sets of well-known test problems (for example, the Rastrigin
function, Ackley function, etc.) were taken as the sets of test functions. The sizes of such sets don't
exceed 100 different functions usually, some of which can be the single-extremal ones (such as the
Rosenbrock function).

@@

\subsection{Parallel Algorithm of Global Search}
\subsubsection{Dimension Reduction with Space-Filling Curves}
Within the framework of the information-statistical global optimization theory,
the Peano space-filling curves (or \textit{evolvents}) \(y(x)\) mapping the interval \([0,1]\)
onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
reduction \cite{sergeyevStronginLera2013, strongin1978, strSergGO}.
\par
As a result of the reduction, the initial multidimensional global optimization
problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
\begin{equation}
\label{eq:oneDimTask}
\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
\end{equation}
\par
It is important to note that this dimensionality reduction scheme transforms the % minimized
Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.
\begin{equation}
\label{eq:holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
\end{equation}
where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem
(\ref{eq:task}).
\par
The algorithms for the numerical construction of the Peano curve approximations are
given in \cite{strSergGO}.

@@

\subsubsection{Algorithm of Global Search on Shared Memory}
\label{sub:ags}
Parallel optimization methods applied in Globalizer to solve the reduced problem
(\ref{eq:oneDimTask}) are based on the MAGS method, which can be presented as follows ---
see \cite{strongin1978}, \cite{strSergGO}.
\par
The initial iteration of the algorithm is performed at an arbitrary point \mbox{\(x^1\in(0,1)\)}.
Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the
following rules.

@@

The convergence conditions and exact formulas for decision rules $R(i)$ and $d(x)$ of the
described algorithm are given, for example, in \cite{strSergGO}.

\subsection{Parallel Algorithm on Distributed Memory Exploiting a Set of Evolvents}
\label{sub:parallel_evolvents}
\subsubsection{Rotated Evolvents}
One of the possible ways to overcome the negative effects of using a numerical
approximation of evolvent (it destroys the information about the neighbor points in
$\mathbb{R}^N$ space)
consists in using the multiple mappings
\begin{equation}
  Y_L(x)=\left\{y^0(x),\ y^1(x),...,\ y^L(x)\right\}
\end{equation}
instead of single Peano curve $y(x)$ (see \cite{strSergGO}).
The building of a set of Peano curves by rotation of the evolvents around the coordinate origin is a
distinctive feature was proposed in \cite{Gergel2009}. Taking into account the initial mapping, one
can conclude that current implementation of the
method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain
onto the corresponding one-dimensional intervals. This method for building a set of mappings can be
``scaled'' easily to obtain more evolvents (up to
$2^N$) if necessary.

>>>>

\documentclass{svproc}
%
% RECOMMENDED
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%
\usepackage{graphicx}
\usepackage{marvosym}
\usepackage{amssymb}
\usepackage{cite}
\usepackage{subfig}
% to typeset URLs, URIs, and DOIs
\usepackage{url}
\usepackage{hyperref}
\usepackage{caption}
\captionsetup[table]{skip=9pt}
\def\UrlFont{\rmfamily}

@@

\begin{document}
\mainmatter              % start of a contribution
%
\title{ComparisonOfSeveralSequentialAndParallelDerivative-freeGlobalOptimizationAlgorithms}
%
\titlerunning{Comparison of several optimization methods}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Vladislav Sovrasov\letter \and Semen Bevzuk}
%
\authorrunning{Vladislav Sovrasov et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Vladislav Sovrasov and Semen Bevzuk}
%
\institute{Lobachevsky State University of Nizhni Novgorod, Russia \\
  \email{sovrasov.vlad@gmail.com, semen.bevzuk@gmail.com}
}

@@

Nonlinear global optimization of non-convex functions is considered to be one of the most
difficult problems of mathematical programming. The finding of the global minimum of
a function of several variables often appears to be more difficult than local optimization in a
thousand-dimensional space. In local optimization applying of the simplest gradient descent
method may appear to be enough, whereas in order to \textit{guarantee} finding the global
optimum, an optimization method has to accumulate information about behavior of the
objective function within the whole search domain
\cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}. Nowadays, various stochastic global
optimization algorithms have attracted much attention, first of all, the evolution ones
\cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}. They have rather simple structure
and allow solving the problems of large dimensionality. However, the evolution algorithms ensure
the global convergence in the probabilistic sense only.

@@

Earlier, the comparison of the stochastic global optimization algorithms \cite{Ali2005, JSSv060i06}
as well as of the deterministic ones \cite{posik2012, Kvasov2018, Liberti2005} between each
other has been considered in the literature. In these works, most of modern methods have
been studied in details, but the emphasis was made mainly on the sequential algorithms while the
reliability and the convergence speed were the main criteria of efficiency of the optimization
methods. In the majority of works, the sets of well-known test problems (for example, the Rastrigin
function, Ackley function, etc.) were taken as the sets of test functions. The sizes of such sets don't
exceed 100 different functions usually, some of which can be the single-extremal ones (such as the
Rosenbrock function).

@@

\subsection{Parallel Algorithm of Global Search}
\subsubsection{Dimension Reduction with Space-Filling Curves}
Within the framework of the information-statistical global optimization theory,
the Peano space-filling curves (or \textit{evolvents}) \(y(x)\) mapping the interval \([0,1]\)
onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
reduction \cite{Sergeyev2013, strongin1978, Strongin2000}.
\par
As a result of the reduction, the initial multidimensional global optimization
problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
\begin{equation}
\label{eq:oneDimTask}
\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
\end{equation}
\par
It is important to note that this dimensionality reduction scheme transforms the % minimized
Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.
\begin{equation}
\label{eq:holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
\end{equation}
where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem
(\ref{eq:task}).
\par
The algorithms for the numerical construction of the Peano curve approximations are
given in \cite{Strongin2000}.

@@

\subsubsection{Algorithm of Global Search on Shared Memory}
\label{sub:ags}
Parallel optimization methods applied in Globalizer to solve the reduced problem
(\ref{eq:oneDimTask}) are based on the MAGS method, which can be presented as follows ---
see \cite{strongin1978}, \cite{Strongin2000}.
\par
The initial iteration of the algorithm is performed at an arbitrary point \mbox{\(x^1\in(0,1)\)}.
Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the
following rules.

@@

The convergence conditions and exact formulas for decision rules $R(i)$ and $d(x)$ of the
described algorithm are given, for example, in \cite{Strongin2000}.

\subsection{Parallel Algorithm on Distributed Memory Exploiting a Set of Evolvents}
\label{sub:parallel_evolvents}
\subsubsection{Rotated Evolvents}
One of the possible ways to overcome the negative effects of using a numerical
approximation of evolvent (it destroys the information about the neighbor points in
$\mathbb{R}^N$ space)
consists in using the multiple mappings
\begin{equation}
  Y_L(x)=\left\{y^0(x),\ y^1(x),...,\ y^L(x)\right\}
\end{equation}
instead of single Peano curve $y(x)$ (see \cite{Strongin2000}).
The building of a set of Peano curves by rotation of the evolvents around the coordinate origin is a
distinctive feature was proposed in \cite{Gergel2009}. Taking into account the initial mapping, one
can conclude that current implementation of the
method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain
onto the corresponding one-dimensional intervals. This method for building a set of mappings can be
``scaled'' easily to obtain more evolvents (up to
$2^N$) if necessary.
